---
title: "Effect of the iPLAY Program on Student Wellbeing"
description: "A Moderation Analysis"
date: "`r Sys.Date()`"
author:
  - first_name: "Taren"
    last_name: "Sanders"
    url: https://www.acu.edu.au/research-and-enterprise/our-research-institutes/institute-for-positive-psychology-and-education/our-people/dr-taren-sanders
    affiliation: Australian Catholic University
    affiliation_url: https://www.acu.edu.au
    orcid_id: 0000-0002-4504-6008
bibliography: References.bib
repository_url: https://github.com/Motivation-and-Behaviour/iPLAY_wellbeing
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source(here::here("packages.R"))
lapply(list.files(here::here("./R"), full.names = TRUE), source)
```

```{r load-targets, include=FALSE}
tar_load(c(desc_table, results_table, plots, df_clean),
  store = here::here("_targets")
)
```
## Introduction

The iPLAY Program is a whole-of-school physical activity intervention, designed to increase student's cardiorespiratory fitness [@lonsdaleScalingupEfficaciousSchoolbased2016].
The program focused on six main areas, which were split between *curricular* and *non-curricular* components.
The curricular components (those implemented by the classroom teacher) are: *Quality PE and School Sport* - providing teachers with the skills and confidence to teach active and engaging PE and school sport lessons; *Classroom Energisers* - encouraging teachers to implement short (less than 5 min) exercise breaks during class time; and *Physically Active Homework* - providing one or more homework activities each week which include physical activity.
The non-curricular components are implemented by up to three *iPLAY Leaders*---teachers seeking leadership roles---in each school, and focus on changes to the school environment.
The components include: *Active Playgroups* - changes to school policy or environment that promote more active break times; *Community Physical Activity Links* - using community sporting groups or providers to provide out-of-hours physical activity opportunities; and *Parents and Caregiver Engagement* - providing parents with information and support to get their children active.

We evaluated iPLAY in two concurrent studies.
In both studies, we collected data from students in Stage 2 (Years 3-4), teachers, and principals before the intervention ('baseline'), 12 months post-baseline (the length of the intervention), and 24 months post-baseline (to test for maintenance of effects).

**Cluster-Randomised Controlled Trial (RCT):** In this study, 22 schools were recruit and matched based on school demographics.
Half of the schools recieved iPLAY, while the other half were waitlisted to recieve iPLAY at the end of the study.
We took measures of children's cardiorespiratory fitness, physical activity levels, school performance, and wellbeing, along with demographic information.

**Dissemination Study:** To evaluate if iPLAY could be delivered 'at scale', we recruited an additional 115 schools who all recieved the iPLAY intervention.
We collected less data from these students, having students complete only a short questionnaire on their self-reported physical activity, and their wellbeing.

In our evaluation of the RCT data [@lonsdaleEffectScalableSchoolBased2021], we found that iPLAY was effective at increasing student's cardiorespiratory fitness.
We also noted a small increase in student's wellbeing post-intervention, but owing to a small sample size, this effect was not statistically significant.
In our evaluation of the dissemination study, we found statistically-significant evidence that iPLAY improved students wellbeing, compared to the students in the RCT control arm.

Given that overall levels of wellbeing across the schools were relatively high, it seems plausible that iPLAY is having a larger benefit for students who have initially low levels of wellbeing.
In this analysis, we test the hypothesis that students with low levels of wellbeing benefit more from the iPLAY intervention than students with high initial wellbeing.

## Method

### Analysis

The outcome variable was students' wellbeing, as reported on the ten-item World Health Organisation's Health Behavior in School-aged Children questionnaire [@robertsHealthBehaviourSchoolaged2009].
We used a mixed-effects model with a gaussian link function.
Mixed-effects model allow us to include random effects for the student and school to account for the clustered study design.
A three-way interaction term (intervention x time point x baseline wellbeing) was used to test for a moderation effects.
Unadjusted and adjusted models were fit, with the latter adjusted for child age, sex,socioeconomic status, indigenous status, and school socioeconomic status and remoteness.
An adjusted model is not stricly required, as the randomisation process should produce balanced covariates across the arms, but was included for the sake of confirming that the randomisation was successful.

## Results

### Sample Description

```{r desc}
desc_stats <-
  df_clean %>%
  dplyr::group_by(time_point) %>%
  summarise(n = n()) %>%
  mutate(n = format(n, big.mark = ","))
```

The sample consisted of `r desc_stats[[1,2]]` students at baseline, `r desc_stats[[2,2]]` at 12 months, and `r desc_stats[[3,2]]` at 24 months.
A description of the sample is provided in Table \@ref(tab:desctable).

```{r desctable}
desc_table
```
### Analysis Results

The result of the adjusted model is shown in Table \@ref(tab:resultstable).
A significant moderation effect of baseline wellbeing was observed at 24 months.

<aside>
You can find the unadjusted model [here](unadjusted_table.html).

A version of the model retricted to those with initial scores of 30 or less is available [here](stratified_table.html).

</aside>

```{r resultstable}
results_table$adjusted
```
```{r savestrat}
gt::gtsave(
  gtsummary::as_gt(results_table$strat),
  here::here("docs", "stratified_table.html")
)
```

```{r saveunadj}
gt::gtsave(
  gtsummary::as_gt(results_table$unadjusted),
  here::here("docs", "unadjusted_table.html")
)
```

To interpret the three-way interaction effect, we plot the group trajectories at different levels of baseline wellbeing.
In this case, we chose four levels to evaluate: students in the 10th (very low wellbeing), 20th (low wellbeing), 50th (i.e., at the median), and 80th (high wellbeing).
These results are available in Figure \@ref(fig:resultsplot).

```{r resultsplot, fig.cap = "Moderation effects of baseline wellbeing on children's wellbeing outcome.", layout="l-page"}
plots
```

## Discussion Points

There's a few interesting things to note here:

* The 'normal' trend for wellbeing is demonstrated in the median trajectory of the control group.
That is, wellbeing tends to be stable for the first 12 months, and then decreases at 24m.
We could hypothesise lots of reasons for this.
For example, children might be more likely to be entering puberty at this stage, or starting to feel the pressure of attending high-school next year.
* There's a steep decline in wellbeing for students with initially high wellbeing.
The control and intervention groups don't differ very much here.
* Students with initially low wellbeing have a steep improvement at 12 months, regardless of if they recieve the intervention or not (this is consistent with the concept of *regression to the mean*).
But, at 24 months there is a strong effect of the intervention on student wellbeing.
* This is consistent with most of the findings from our previous research.
Even though iPLAY was a 12 month intervention, it seems that it took longer than this for the effects on students to appear.